Untitled

### 2020-11-10 ms/manuscript.Rmd
Update MS

---
title: ""
bibliography: references.bib
mainfont: Roboto-Regular.ttf
fontsize: 12pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: none # Use pandoc default, otherwise csl can't be used for reference formatting
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: yes
    latex_engine: xelatex
    pandoc_args: [ "--csl", "journal-of-plant-research.csl"]
    includes:
      in_header: journal-of-plant-research.sty
header-includes: 
  \usepackage{float}
  \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
editor_options: 
  markdown: 
    wrap: sentence
  chunk_output_type: console
---

@@

```{r load-data}
loadd(
  list = c(
    "climate"
  )
)
```

@@

<!-- See https://stackoverflow.com/questions/24672111/how-to-add-a-page-break-in-word-document-generated-by-rstudio-markdown/40474753
for inserting page breaks when rendering to MS word -->
#####

\newpage

@@

#####

\newpage

@@

We selected the island of Moorea, French Polynesia (17°32' S, 149°50' W) for our survey of filmy fern ecophysiology (Fig. 2).
Moorea is a small (135 km^2^; maximum elevation 1206 m) tropical oceanic island, located ca.
15 km NW of its larger sister island, Tahiti.
Despite its small size, it hosts a range of habitats including coastal strand, low elevation rainforest, and high elevation cloud forest.
At the time of the study, a total of 18 filmy fern species were known from Moorea [@Murdock2003; @Ranker2005; @Nitta2011b; @Florence], including multiple terrestrial, saxicolous, and epiphytic species.
The two major clades of filmy ferns, the hymenophylloid and trichomanoid lineages [@Pryer2001a], are represented on Moorea by similar numbers of species.
We follow the taxonomic system of @Ebihara2006.

@@

We surveyed filmy ferns on Moorea as part of a larger study including all ferns of Moorea and Tahiti [@Nitta2017].
During this study, 17 sampling sites were established from \circa{ }200 m to 1200 m primarily along trails to the summits of three large peaks on Moorea: Mt. Rotui (899 m), Mt. Mouaputa (880 m), and Mt. Tohiea (1206 m; Fig. 2).
At each site, sporophytes were sampled in a single 10 m × 10 m plot, and gametophytes sampled in two to three 50 cm × 50 cm subplots located within the sporophyte plot.
Epiphytes were sampled to a maximum height of \circa{ }2 m on their host plants.
Unlike many other fern species, filmy fern gametophytes have unique morphologies that allow them to be identified to family [\eg, filamentous, ribbon with round gemmae; @Nayar1971].
This allowed us to specifically use field-collected filmy fern gametophytes for physiological tests, which we later identified to species using DNA barcoding (see below).
We scored growth habit of each species based on field observations as terrestrial, saxicolous, low elevation epiphyte (mostly occurring outside of cloud forest, \ie, below \circa{ }500 m), or high elevation epiphyte (mostly occurring in cloud forest, \ie, above \circa{ }500 m).

@@

Twenty-two filmy fern taxa (including one new record and two species that were recognized as multiple varieties) were observed as sporophytes, and 18 as gametophytes on Moorea in total (Table 1).
*Hymenophyllum braithwaitei* (Ebihara & K. Iwats.), only observed near the peak of Mt. Tohiea (1206 m), is a new record for Moorea and French Polynesia.
It was previously known only from Vanuatu and New Caledonia (Ebihara et al., 2010).
Three taxa with distinct *rbcL* sequences, morphology, and elevational ranges could be distinguished within the *Crepidomanes minutum* (Blume) K. Iwats. species complex, but species status of these taxa remains uncertain (Nitta et al., 2011a).
We refer to these here as *C. minutum* var. 1, 2, and 3.
Similarly, two members of the *Abrodictyum asae-grayi* (Bosch) Ebihara & K. Iwats. complex (also including *Abrodictyum caudatum* (Brack.) Ebihara & K. Iwats. in our tree; Appendix S1) were observed on the basis of *rbcL* and morphology, but as taxonomic treatment is beyond the scope of the current study we refer to them as *A. asae-grayi* var. 1 and 2.
In several species, gametophytes exceeded the maximum or minimum elevational range of sporophytes (Table 1) including *Callistopteris apiifolia* (C. Presl) Copel., *Crepidomanes minutum* var. 2 and 3, *Hymenophyllum digitatum* (Sw.) Fosberg, *Hymenophyllum javanicum* Spreng., *Hymenophyllum pallidum* (Blume) Ebihara & K.
Iwats., and *Polyphlebium borbonicum* (Bosch) Ebihara & Dubuisson.
All of these have sporophytes that are normally only found in cloud forest (ca. 500--1200 m), but gametophytes that are distributed to lower elevations.

@@

Mean daily temperature was strongly correlated with elevation, and dropped ca.
6 *◦*C from the lowest surveyed site (201 m) to the highest (1206 m; Fig. 3).
Although epiphytic and terrestrial habitats were similar in temperature, epiphytic habitats tended to have lower daily mean RH, and greater daily variation in RH, than terrestrial ones (Fig. 3, Appendix S2).
This difference was greatest at low elevations, and smallest at high elevations.
One significant departure from these trends was observed at the Mt. Rotui 830 m site (indicated by arrows in Fig. 3).
This site is much more exposed than the other cloud forest sites (Fig. 2), and many of the plants on the N facing slope are stunted (J. Nitta, pers. obs.).
Daily variation in temperature and RH was much greater, and mean RH much lower, at this site than other high elevation sites.

@@

Results of the tests for phylogenetic signal (Blomberg's *K* and Pagel's λ*)* were similar for sporophytes and gametophytes, revealing significant phylogenetic signal for DT (gametophytes *K* and λ both near 1, *P \<*0.05; sporophytes *K* = 0.78, *P* = 0.02 and λ = 0.96, *P* = 0.11), but no signal for PPFD~95%~ or ETR~max~ (Table 2).

>>>>

---
title: ""
bibliography: 
  - references.bib
  - references_other.yaml
mainfont: Roboto-Regular.ttf
fontsize: 12pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: none # Use pandoc default, otherwise csl can't be used for reference formatting
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: yes
    latex_engine: xelatex
    pandoc_args: [ "--csl", "journal-of-plant-research.csl"]
    includes:
      in_header: journal-of-plant-research.sty
header-includes: 
  \usepackage{float}
  \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
editor_options: 
  markdown: 
    wrap: sentence
  chunk_output_type: console
---

@@

```{r load-data}
loadd(
  list = c(
    "climate",
    "filmy_species",
    "moorea_sites",
    "phylosig",
    "range_types"
  )
)
```

@@

\clearpage
#####

@@

\clearpage
#####

@@

```{r study-system-stats}
# Calculate number of all filmy fern species, trichomanoids, and hymenophylloids
# (not considering numbered taxa)

n_filmy_sp <- str_remove_all(filmy_species, "[0-9]") %>% 
  n_distinct()

n_trich_sp <- str_remove_all(filmy_species, "[0-9]") %>% 
  magrittr::extract(., str_detect(., "^Hymeno", negate = TRUE)) %>%
  n_distinct()

n_hymeno_sp <- str_remove_all(filmy_species, "[0-9]") %>% 
  magrittr::extract(., str_detect(., "^Hymeno", negate = FALSE)) %>%
  n_distinct()

```

We selected the island of Moorea, French Polynesia (17°32' S, 149°50' W) for our survey of filmy fern ecophysiology (Fig. 2).
Moorea is a small (135 km^2^; maximum elevation 1206 m) tropical oceanic island, located ca.
15 km NW of its larger sister island, Tahiti.
Despite its small size, it hosts a range of habitats including coastal strand, low elevation rainforest, and high elevation cloud forest.
A total of `r n_filmy_sp` filmy fern species are known from Moorea, including multiple terrestrial, saxicolous, and epiphytic species [@Murdock2003; @Ranker2005; @Nitta2011b; @Nitta2017; @Florence].
The two major clades of filmy ferns, the hymenophylloid and trichomanoid lineages [@Pryer2001a], are represented on Moorea by `r n_hymeno_sp` and `r n_trich_sp` species, respectively.
We follow the taxonomic system of @Ebihara2006.

@@

We surveyed filmy ferns on Moorea as part of a larger study including all ferns of Moorea and Tahiti [@Nitta2017].
During this study, `r n_distinct(moorea_sites$site)` sampling sites were established from \circa{ }200 m to 1200 m primarily along trails to the summits of three large peaks on Moorea: Mt. Rotui (899 m), Mt. Mouaputa (880 m), and Mt. Tohiea (1206 m; Fig. 2).
At each site, sporophytes were sampled in a single 10 m × 10 m plot, and gametophytes sampled in two to three 50 cm × 50 cm subplots located within the sporophyte plot.
Epiphytes were sampled to a maximum height of \circa{ }2 m on their host plants.
Unlike many other fern species, filmy fern gametophytes have unique morphologies that allow them to be identified to family [\eg, filamentous, ribbon with round gemmae; @Nayar1971].
This allowed us to specifically use field-collected filmy fern gametophytes for physiological tests, which we later identified to species using DNA barcoding (see below).
We scored growth habit of each species based on field observations as terrestrial, saxicolous, low elevation epiphyte (mostly occurring outside of cloud forest, \ie, below \circa{ }500 m), or high elevation epiphyte (mostly occurring in cloud forest, \ie, above \circa{ }500 m).

@@

In total, `r n_distinct(range_types$species)` filmy fern taxa (including two species that were recognized as multiple varieties) were observed on Moorea.
Three taxa with distinct *rbcL* sequences, morphology, and elevational ranges could be distinguished within the *Crepidomanes minutum* (Blume) K. Iwats. species complex, but species status of these taxa remains uncertain (Nitta et al., 2011a).
We refer to these here as *C. minutum* var. 1, 2, and 3.
Similarly, two members of the *Abrodictyum asae-grayi* (Bosch) Ebihara & K. Iwats. complex (also including *Abrodictyum caudatum* (Brack.) Ebihara & K. Iwats. in our tree; Appendix S1) were observed on the basis of *rbcL* and morphology, but as taxonomic treatment is beyond the scope of the current study we refer to them as *A. asae-grayi* var. 1 and 2.

In several species, gametophytes exceeded the maximum or minimum elevational range of sporophytes (Table 1) including *Callistopteris apiifolia* (C. Presl) Copel., *Crepidomanes minutum* var. 2 and 3, *Hymenophyllum digitatum* (Sw.) Fosberg, *Hymenophyllum javanicum* Spreng., *Hymenophyllum pallidum* (Blume) Ebihara & K.
Iwats., and *Polyphlebium borbonicum* (Bosch) Ebihara & Dubuisson.
All of these have sporophytes that are normally only found in cloud forest (ca. 500--1200 m), but gametophytes that are distributed to lower elevations.

Sporophytes of two taxa, *C. minutum* var. 3 and *Hymenophyllum braithwaitei* (Ebihara & K. Iwats.), did not occur in the study plots, but each was observed from a single population on Mt. Tohiea, at ca. 1000 m and near the summit (1206 m), respectively.
We include these observations for comparison of ranges between sporophytes and gametophytes.

@@

Mean daily temperature was strongly correlated with elevation, and dropped ca.
6 °C from the lowest surveyed site (201 m) to the highest (1206 m; Fig. 3).
Although epiphytic and terrestrial habitats were similar in temperature, epiphytic habitats tended to have lower daily mean RH, and greater daily variation in RH, than terrestrial ones (Fig. 3, Appendix S2).
This difference was greatest at low elevations, and smallest at high elevations.
One significant departure from these trends was observed at the Mt. Rotui 830 m site (indicated by arrows in Fig. 3).
This site is much more exposed than the other cloud forest sites (Fig. 2), and many of the plants on the N facing slope are stunted (J. Nitta, pers. obs.).
Daily variation in temperature and RH was much greater, and mean RH much lower, at this site than other high elevation sites.

@@

```{r phy-sig-stats}
sporo_recover_k <- phylosig %>% filter(generation == "sporo", trait == "recover_mean") %>% pull(kval) %>% jntools::round_t(2)
sporo_recover_k_pval <- phylosig %>% filter(generation == "sporo", trait == "recover_mean") %>% pull(k_pval) %>% jntools::round_t(3)
sporo_recover_lambda <- phylosig %>% filter(generation == "sporo", trait == "recover_mean") %>% pull(lambda) %>% jntools::round_t(2)
sporo_recover_lambda_pval <- phylosig %>% filter(generation == "sporo", trait == "recover_mean") %>% pull(lambda_pval) %>% jntools::round_t(3)

# Make sure that all ETRmax and PAR have pval > 0.05
phylosig %>% 
  filter(trait != "recover_mean") %>%
  verify(all(k_pval > 0.05)) %>%
  verify(all(lambda_pval > 0.05), success_fun = success_logical, error_fun = error_stop)
```

Results of the tests for phylogenetic signal (Blomberg's *K* and Pagel's λ*)* were similar for sporophytes and gametophytes, revealing significant phylogenetic signal for DT (gametophytes *K* and λ both near 1, *P \<*0.05; sporophytes *K* = `r sporo_recover_k`, *P* = `r sporo_recover_k_pval` and λ = `r sporo_recover_lambda`, *P* = `r sporo_recover_lambda_pval`), but no signal for PPFD~95%~ or ETR~max~ (Table 2).
